### 2015-10-14 papers/announcement/src/parallel/parallel.tex

\section{Benchmarks on parallel frameworks}\label{sec:parallel}

{\color{red} To be written.}

\subsection{OpenMP benchmark}\label{subsec:openmp}

{\color{red} To be written.}

\subsection{MPI benchmark}\label{subsec:mpi}

{\color{red} To be written.}