### 2015-10-16 papers/announcement/src/parallel/openmp.tex
improve openmp benchmark

\subsection{OpenMP benchmark}\label{subsec:openmp}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.95\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling.png}
    \caption{Strong scaling, number of cells 5000}\label{fig:strong-scaling-openmp}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.95\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling.png}
    \caption{Weak scaling, minimum number of cells 500}\label{fig:weak-scaling-openmp}
  \end{subfigure}\\
  \caption{Scaling efficiency with OpenMP}\label{fig:scaling-openmp}
\end{figure}