### 2015-10-16 papers/announcement/src/parallel/parallel.tex

\section{Benchmarks on parallel frameworks}\label{sec:parallel}

Speedup of strong scaling:

\begin{equation}
  speedup(N, k) = \frac{T_{serial}(N)}{T_{parallel}(N, k)}
  \label{eq:strong-scaling-speedup}
\end{equation}
where $N$ is the problem size, $K$ the number of cpu-cores used, $T_{serial}$ is the cpu time of the serial code and $T_{parallel}$ the one of the parallel code.

Efficiency of strong scaling:

\begin{equation}
  efficiency(N, k) = \frac{speedup(N, k)}{k}
  \label{eq:strong-scaling-efficiency}
\end{equation}

Sizeup of weak scaling:

\begin{equation}
  sizeup(N, k) = \frac{N_k}{N_1} \cdot \frac{T_{serial}(N_1)}{T_{parallel}(N_k, k)}
  \label{eq:weak-scaling-sizeup}
\end{equation}
where $N_k=k \cdot N_1$.

Efficiency of weak scaling:

\begin{equation}
  efficiency(N, k) = \frac{sizeup(N, k)}{k}
  \label{eq:weak-scaling-efficiency}
\end{equation}

{\color{red} To be written.}

\input{src/parallel/openmp.tex}

\subsection{MPI benchmark}\label{subsec:mpi}

{\color{red} To be written.}